---
title: "Exploratory Data Analysis"
author: "Critical Thinking Group 3"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
---


```{r knitr_init, echo=FALSE, cache=FALSE, message=FALSE}
library(knitr)
library(rmdformats)
library(dplyr)
library(tidyr)
library(ggplot2)
library(kableExtra)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

```{r read_data}
training <- read.csv("../data/moneyball-training-data.csv") %>%
  #select(-TEAM_BATTING_HBP, -INDEX) %>%
  mutate(TEAM_BATTING_S = TEAM_BATTING_H + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR)
```

## Scatter Plots

```{r small_multiples_scatter, warning=FALSE}
training %>%
  gather(variable, value, TEAM_BATTING_H:TEAM_BATTING_S) %>%
  ggplot(., aes(TARGET_WINS, value)) + 
  geom_point(fill = "indianred4", color="indianred4") + 
  facet_wrap(~variable, scales ="free", ncol = 4) +
  labs(x = element_blank(), y = "Wins")
```


```{r}
training %>%
  summary() %>%
  kable() %>%
  kable_styling()
```

## Distribution Plots

```{r small_multiples_density, warning=FALSE}
training %>%
  gather(variable, value, TARGET_WINS:TEAM_BATTING_S) %>%
  ggplot(., aes(value)) + 
  geom_density(fill = "indianred4", color="indianred4") + 
  facet_wrap(~variable, scales ="free", ncol = 4) +
  labs(x = element_blank(), y = element_blank())
```

## Preprocessing

### Data Transformations

```{r}
transformations <- function(df){
  right_skewed_variables <- c("TEAM_BASERUN_CS", "TEAM_BASERUN_SB", "TEAM_BATTING_3B")
  for(v in right_skewed_variables){
    # Apply a square root transformation
    training[[v]] <- sqrt(training[[v]])
  }
  df
}

training <- training %>% transformations()
```

```{r small_multiples_transformed_density, warning=FALSE}
training %>%
  gather(variable, value, TARGET_WINS:TEAM_BATTING_S) %>%
  ggplot(., aes(value)) + 
  geom_density(fill = "indianred4", color="indianred4") + 
  facet_wrap(~variable, scales ="free", ncol = 4) +
  labs(x = element_blank(), y = element_blank())
```

```{r}
training %>%
  summary() %>%
  kable() %>%
  kable_styling()
```

### Remove Outliers

```{r}
predictors <- training %>% select(-TARGET_WINS, -INDEX) %>% names()
for(v in predictors){
#for (v in names(training)){
    outliers <- boxplot(training[[v]], plot = FALSE)$out
    if(length(outliers) > 0){
      if(!exists("the_outliers")){
        #the_outliers <- training[which(training[[v]] %in% outliers),]
      } else {
        #the_outliers <- rbind(the_outliers, training[which(training[[v]] %in% outliers),])
      }
      training <- training[-which(training[[v]] %in% outliers),]
      message(paste("Removing", length(outliers), v, "outliers"))
    }
}
```

```{r small_multiples_outliers_removed, warning=FALSE}
training %>%
  gather(variable, value, TARGET_WINS:TEAM_BATTING_S) %>%
  ggplot(., aes(value)) + 
  geom_density(fill = "indianred4", color="indianred4") + 
  facet_wrap(~variable, scales ="free", ncol = 4) +
  labs(x = element_blank(), y = element_blank())
```

```{r}
training %>%
  summary() %>%
  kable() %>%
  kable_styling()
```


```{r}
adjust_by_a_factor_of <- .1
adjust_by <- c()
adjust_by_names <- c()
for (v in names(training)[2:16]){
  temp <- median((training[[v]] * adjust_by_a_factor_of) / training$TARGET_WINS, na.rm = TRUE)
  adjust_by <- append(adjust_by, temp)
  adjust_by_names <- append(adjust_by_names, v)
}

names(adjust_by) <- adjust_by_names
```



```{r small_multiples_adjusted_scatter, warning=FALSE}
training %>%
  gather(variable, value, TEAM_BATTING_H:TEAM_BATTING_S) %>%
  ggplot(., aes(TARGET_WINS, value)) + 
  geom_point(fill = "indianred4", color="indianred4") + 
  facet_wrap(~variable, scales ="free", ncol = 4) +
  labs(x = element_blank(), y = "Wins")
```


## Interesting

Groups emerged from the predictors:

```{r}
training <- training %>%
  mutate(ratio = TEAM_BATTING_H / TEAM_PITCHING_H,
         Group = ifelse(ratio > 0.98, "A", "B")) %>%
  mutate(Group = ifelse(ratio < 0.9, "C", Group)) %>%
  mutate(Group = as.factor(Group))

training %>%
  ggplot(., aes(TEAM_BATTING_H, TEAM_PITCHING_H, color = Group)) +
         geom_point()
```

They don't have much meaning though

```{r}
ggplot(training, aes(TEAM_BATTING_H, TARGET_WINS, color = Group)) + 
  geom_point()
```


This one is not very useful either

```{r}
training <- training %>%
  mutate(ratio = TEAM_BATTING_BB / TEAM_PITCHING_BB,
         Group = ifelse(ratio > 0.98, "A", "B")) %>%
  mutate(Group = ifelse(ratio < 0.9, "C", Group)) %>%
  mutate(Group = as.factor(Group))

training %>%
  ggplot(., aes(TEAM_BATTING_BB, TEAM_PITCHING_BB, color = Group)) +
         geom_point()

ggplot(training, aes(TEAM_PITCHING_BB, TARGET_WINS, color = Group)) + 
  geom_point()
```

Same story with this one

```{r}
training <- training %>%
  mutate(ratio = TEAM_BATTING_HR / TEAM_PITCHING_HR,
         Group = ifelse(ratio > 0.98, "A", "B")) %>%
  mutate(Group = ifelse(ratio < 0.9, "C", Group)) %>%
  mutate(Group = as.factor(Group))

training %>%
  ggplot(., aes(TEAM_BATTING_BB, TEAM_PITCHING_BB, color = Group)) +
         geom_point()

ggplot(training, aes(TEAM_PITCHING_BB, TARGET_WINS, color = Group)) + 
  geom_point()
```


But this one is very interesting because it seperates the bimodal distribution.

```{r}
training <- training %>%
  mutate(ratio = TEAM_BATTING_SO / TEAM_PITCHING_SO,
         Group = ifelse(ratio > 0.98, "A", "B")) %>%
  mutate(Group = ifelse(ratio < 0.9, "C", Group)) %>%
  mutate(Group = as.factor(Group))

training %>%
  ggplot(., aes(TEAM_BATTING_H, TEAM_PITCHING_H, color = Group)) +
         geom_point()

ggplot(training, aes(TEAM_PITCHING_SO, TARGET_WINS, color = Group)) + 
  geom_point()

ggplot(training, aes(TEAM_PITCHING_SO, color = Group)) + 
  geom_density()
```


This only happens after the outliers are pulled out.



```{r eval=FALSE}
# for(x in predictors){
#   for(y in predictors){
#     if(x != y){
#       print(paste(x,"&",y))
#       plot(training[[x]], training[[y]])
#     }
#   }
# }
```